---
title: "World Happiness report 2016"
author: "Maria Carolina da S. Andrea"
date: "9 de abril de 2020"
output: html_document
---

<br></br>

#### Happiness rank and scores by country, 2016.

This data is the World Happiness Report.
I obtained this from Kaggle website: https://www.kaggle.com/unsdsn/world-happiness

This data refers to the year of 2016.
Details about the data, what it represents can be obtained in Kaggle's webpage.


<br></br>

```{r setup, include=FALSE, message=FALSE, warning=FALSE}

# importing data, renaming columns

library(tidyverse)

h_2016 <- read.csv("h_2016.csv", header = TRUE)

head(h_2016)

colnames(h_2016) <- c('Country', 'Region', 'H_Rank', 'H_Score',
                      'LCI', 'UCI', 'Gdp_pc', 'Family', 'Life_ex',
                      'Freedom', 'Corrup_ppt', 'Generosity', 'Dystopia_res')

# Normalize the scores

library(caret)


h_2016_norm <-
    h_2016 %>%
    mutate_at(scale, .vars = vars(-c(1:6,13)))

head(h_2016_norm)


# check correlations

cor_1 <- cor(h_2016_norm[,-c(1,2)])


cor_2 <- cor(h_2016_norm[,-c(1,2,3,5,6)])
cor_2

```

<br></br>

## Happiness-related scores correlation
<br></br>

### Worldwide

```{r echo=FALSE, message=FALSE, warning=FALSE}

library(corrgram)

corrgram(cor_2, order=T, lower.panel=panel.shade,
         upper.panel=panel.cor, text.panel=panel.txt,
         main = 'Happiness-related scores Worldwide')

```

It is interesting to notice that apparently, Happiness score was **more strongly and positively related to GDP per capita, Life expectancy and Family** (all above 0.70),less strongly related to Freedom and perception of corruption. And the **least correlated score was Generosity**.

However, this pattern might change among world's main regions.
Since there is a lot of countries, I investigated the 10 Regions given
by the dataset.

Although slightly tedious, results show some interesting differences among the regions.

Let's check one correlogram per each of the 10 regions.

<br></br>

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Eastern Asia
easia <- subset(h_2016_norm, Region == "Eastern Asia")
cor_easia <- cor(easia[,-c(1,2,3,5,6)])

corrgram(cor_easia, order=T, lower.panel=panel.shade,
         upper.panel=panel.cor, text.panel=panel.txt,
         main = 'Happiness-related scores in Eastern Asia')

```

In Eastern Asia, GDP per capita stands out among the other scores, considering the
more positive correlation.
Freedom, Family and Generosity are weakly correlated (below 0%).

Let's check other parts of Asia.

<br></br>

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Southeastern Asia
soeas <- subset(h_2016_norm, Region == "Southeastern Asia")
cor_soeas <- cor(soeas[,-c(1,2,3,5,6)])

corrgram(cor_soeas, order=T, lower.panel=panel.shade,
         upper.panel=panel.cor, text.panel=panel.txt,
         main = 'Happiness-related scores in Southeastern Asia')

```

Also, GDP plays the most important role in correlating to Happiness Score.
Freedom and Generosity are weakly correlated (below 0%).

<br></br>

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Southern Asia
srtas <- subset(h_2016_norm, Region == "Southern Asia")
cor_srtas <- cor(srtas[,-c(1,2,3,5,6)])

corrgram(cor_srtas, order=T, lower.panel=panel.shade,
         upper.panel=panel.cor, text.panel=panel.txt,
         main = 'Happiness-related scores in Southern Asia')

```

In Southern Asia, interestingly, Corruption perception is the most related to Happines Score.
Freedom and Generosity are weakly correlated (below 10%).

<br></br>

Now moving to Africa.

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Middle East and Northern Africa
menaf <- subset(h_2016_norm, Region == "Middle East and Northern Africa")
cor_menaf <- cor(menaf[,-c(1,2,3,5,6)])

corrgram(cor_menaf, order=T, lower.panel=panel.shade,
         upper.panel=panel.cor, text.panel=panel.txt,
         main = 'Happiness-related scores in Middle East and Northern Africa')

```
Now for the first time in this analysis, Family appear as the most correlated. GDP comes right after.
Generosity is the least correlated (below 0).

<br></br>


```{r echo=FALSE, message=FALSE, warning=FALSE}

# Sub-Saharan Africa
ssafr <- subset(h_2016_norm, Region == "Sub-Saharan Africa")
cor_ssafr <- cor(ssafr[,-c(1,2,3,5,6)])

corrgram(cor_ssafr, order=T, lower.panel=panel.shade,
         upper.panel=panel.cor, text.panel=panel.txt,
         main = 'Happiness-related scores in Sub-Saharan Africa')

```

None of the factor present stronger correlation values (above 50%).
Freedom presented the strongest correlation, with Family and GDP right after.
Corruption perception was the least correlated to Happiness score.

<br></br>

Now Southern portion of America.

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Latin America and Caribbean
lamca <- subset(h_2016_norm, Region == "Latin America and Caribbean")
cor_lamca <- cor(lamca[,-c(1,2,3,5,6)])

corrgram(cor_lamca, order=T, lower.panel=panel.shade,
         upper.panel=panel.cor, text.panel=panel.txt,
         main = 'Happiness-related scores in Latin America and Caribbean')

```

In Southern America, GDP also appears with the strongest correlation.
Health appears right after.
Generosity was the least correlated (below 0).


<br></br>

Moving to Europe.

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Western Europe
wseur <- subset(h_2016_norm, Region == "Western Europe")
cor_wseur <- cor(wseur[,-c(1,2,3,5,6)])

corrgram(cor_wseur, order=T, lower.panel=panel.shade,
         upper.panel=panel.cor, text.panel=panel.txt,
         main = 'Happiness-related scores in Western Europe')

```

Corruption perception was the strongest correlation, and very close to Family and Freedom.
Health-related factor was the weakest correlation (almost 0).

<br></br>


```{r echo=FALSE, message=FALSE, warning=FALSE}

# Central and Eastern Europe
ceeur <- subset(h_2016_norm, Region == "Central and Eastern Europe")
cor_ceeur <- cor(ceeur[,-c(1,2,3,5,6)])

corrgram(cor_ceeur, order=T, lower.panel=panel.shade,
         upper.panel=panel.cor, text.panel=panel.txt,
         main = 'Happiness-related scores in Central and Eastern Europe')

```

In Eastern Europe, the strongest correlation was with Family.
Health, generosity and corruption perception were the weakest correlations.

<br></br>

Now parts of Oceania.

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Australia and New Zealand
ausnz <- subset(h_2016_norm, Region == "Australia and New Zealand")
cor_ausnz <- cor(ausnz[,-c(1,2,3,5,6)])

corrgram(cor_ausnz, order=T, lower.panel=panel.shade,
         upper.panel=panel.cor, text.panel=panel.txt,
         main = 'Happiness-related scores in Australia and New Zealand')

```

<br></br>

And finally the northern portion of America.

```{r echo=FALSE, message=FALSE, warning=FALSE}

# North America
noamr <- subset(h_2016_norm, Region == "North America")
cor_noamr <- cor(noamr[,-c(1,2,3,5,6)])

corrgram(cor_noamr, order=T, lower.panel=panel.shade,
         upper.panel=panel.cor, text.panel=panel.txt,
         main = 'Happiness-related scores in North America')

```
<br></br>
    
Correlograms are weird for Oceania countries and North America.
That may be relate to these subsets of data having only two countries in each.

But these correlograms helped to begin to see some differences among the countries.


<br></br>

#### Grouping the data

##### Cluster Analysis


In order to "divide" more the countries and regions looking for some
similarities, I'll use a partitioning clustering.

Using K-means clustering and finding the best number of clusters, it is
possible to look at countries and regions in a slightly different manner.

<br></br>

```{r echo=FALSE, message=FALSE, warning=FALSE}

# i'll create a df with the regions as rownames

df_to_cluster <- h_2016_norm[,-c(2,3,5,6)]
row.names(df_to_cluster) <- df_to_cluster$Country
df_to_cluster[1] <- NULL


# number of clusters and k-means clustering

library(factoextra)
library(ggsci)

fviz_nbclust(df_to_cluster, kmeans,
             method = "gap_stat")

km.res <- kmeans(df_to_cluster, 3, nstart = 25)
```


```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=12}

# Visualize
fviz_cluster(km.res, data = df_to_cluster,
             ellipse.type = "convex",
             palette = "nrc",
             repel = TRUE,
             ggtheme = theme_minimal())

```

<br></br>

So, working with the recommended 3 clusters, it is normal to wonder wether
regions' countries fall entirely (or at least its majority) into each cluster.

Let's investigate regions and countries within each cluster.

<br></br>

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=12}

# transform cluster results into a df
df <- as.data.frame(km.res$cluster)

# return rownmes to column
df <-
    df %>%
    rownames_to_column(var = "Country")

# join these to add the cluster assignment into my main df

h_2016_cl <-
    h_2016_norm %>%
    left_join(df)

h_2016_cl %>%
    arrange(desc(H_Score)) %>%
    ggplot(aes(x = Country, y = H_Rank, size = H_Score, fill = Region))+
    geom_point(alpha = 0.6, shape = 21, color = 'black')+
    scale_size(range = c(.1, 20), name = 'Happiness Score')+
    facet_grid(.~km.res$cluster, scales = 'free_x')+
    #scale_y_continuous(limits=c(157,0))+
    scale_fill_npg() +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 70, vjust = 0.5))+
    theme(legend.position="bottom")+
    guides(fill=guide_legend(nrow=4,byrow=TRUE))+
    ylab("Happiness Rank")+
    xlab('Country')
    #xlab("Gdp per Capita")

```

Although not quite there yet, we can also get some useful insights from this plot.

Cluster 1 has countries with the highest Happines Ranked countries, and is also the cluster with less countries.

Cluster 2 apparentely has the greater amount of countries and a variety of regions.
Here, the Happiness Rank of countries shows the higher dispersion: countries in the top ranked, and countries at the bottom of the rank.

Cluster 3 has countries wich appear as the lowest ranked in the Happines Rank.

But is still difficult to say which regions are in which clusters.

<br></br>

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=12}

# number and proportion of clusters to total countries
cn_per_cl <- as.data.frame(km.res$size)

# proportion (related to total number of countries)
pr_per_cl <- as.data.frame(km.res$size/157 * 100)

# number of countries per region
# (use this to the porportion of countries' region in each cluster)
c_per_rg <-
    h_2016_cl %>%
    group_by(Region)%>%
    summarise(n_ct_per_rg = n())

# number of countries (and their regions) within each cluster
c_per_cl <-
    h_2016_cl %>%
    group_by(Region, km.res$cluster) %>%
    # number of countries of regions within each cluster
    summarise(n_ctreg_per_cl = n())

# and finally, for each cluster, the proportion of a region's countries in it!
cl_reg_prop <-
    c_per_cl %>%
    left_join(c_per_rg) %>%
    mutate(proportion = (n_ctreg_per_cl / n_ct_per_rg) *100)

# Now let's visualize this
cl_reg_prop$proportion <- round(cl_reg_prop$proportion, digits = 2)

cl_reg_prop$Region <- factor(cl_reg_prop$Region,
                             levels = c('Australia and New Zealand',
                                        'North America', 'Western Europe',
                                        'Central and Eastern Europe',
                                        'Latin America and Caribbean',
                                        'Eastern Asia', 'Middle East and Northern Africa',
                                        'Southeastern Asia', 'Southern Asia',
                                        'Sub-Saharan Africa'))

cl_reg_prop %>%
    ggplot(aes(x = Region, y = proportion, fill = as.factor(`km.res$cluster`), label = proportion))+
    geom_col(color = 'black', width = .7)+
    geom_text(size = 3, position = position_stack(vjust = 0.5))+
    scale_fill_npg(name = 'Cluster')+
    theme_classic() +
    theme(axis.text.x = element_text(angle = 70, vjust = 0.5))+
    xlab("Main world region")+
    ylab("Proportion of countries in clusters across regions (%)")


```


<br></br>

(Partial) conclusions.

Since a regression analysis within a year of data does not seem to be a coherent option, I focused on exploring the data and trying to understand differences among regions.

It became clear that main world regions are different among each other (of course, even countries are different among each other), and they perceive and value happiness in different manners.

Each country's history probably play a role in this.

Fewer countries appear as the highest on the Happines-Rank.
These are also, majoritarily, the richest countries (North America, Australia, New Zealand,
Western Europe).

The lowest-ranked countries in terms of happiness are formed in its majority by the poorest countries.

I intend to continue the analysis, with further statistics comparing years.
